Semi-supervised author disambiguation of scientific publications

### 2015-08-05 paper.tex
save

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem, digital
libraries are increasingly offering tools for authors to manually curate their
publications and claim which ones are theirs. Indirectly, these tools allow for
the inexpensive collection of large annotated training data, which can be
further leveraged to build an automated disambiguation system capable of
inferring features or patterns for identifying publications written by a same
person.  In this context, we propose a semi-supervised machine learning
disambiguation solution in the case large and partially annotated databases of
scientific publications. Using a real-world dataset of 1 million manual
annotations, we demonstrate how we exploited this data (i) to learn an accurate
classifier for identifying corefering authors, (ii) to guide the clustering of
scientific publications by distinct authors in a semi-supervised way and
finally (iii) to match the resulting clusters with existing profiles.  Our
analysis of parameters and features on this large training set reveal that...
\glnote{To be completed with actual results.}

@@

In academic digital libraries, author name disambiguation is the problem of
grouping together publications written by a same person.  With the fast growth
of the scientific literature, this has become a pressing issue because the
accuracy of information managed at the level of individuals directly affects
the relevance search results (e.g., when querying for all publications written
by a given author), the reliability of bibliometrics (e.g., citation counts or
other impact metrics) or the relevance of scientific network analysis. Author
name disambiguation is often a difficult problem because an author may use
different spellings, variants or synonyms across her career and/or distinct
authors may share the same name. In particular, author disambiguation becomes
even more critical for researchers from non-Western cultures, where personal names
may be traditionally less diverse (leading to homonym issues) or for which
transliteration to latin characters may not be unique (leading to synonym
issues).

@@

Formally, let us assume a set of publications ${\cal P} = \{ p_0, ...,
p_{N-1}\}$ along with the set of unique individuals ${\cal A} = \{ a_0, ...,
a_{M-1}\}$ having together authored all publications in ${\cal P}$.  Let us
define a signature $s \in p$ from a publication as a unique piece of
information identifying one of the authors of $p$ (e.g., the author name, his
affiliation, along with any other metadata can be derived from $p$). Let us
denote by ${\cal S} = \{ s | s \in p, p \in {\cal P} \}$ the set of all
signatures that can be extracted from all publications in ${\cal P}$. In this
framework, author disambiguation can be stated as the problem of
finding a partition ${\cal C} = \{ c_0, ..., c_{M-1} \}$ of ${\cal S}$ such
that ${\cal S} = \cup_{i=0}^{M-1} c_i$, $c_i \cap c_j = \phi$ for all $i \neq
j$, and where subsets $c_i$, or clusters, each corresponds to the set of all
signatures belonging to the same individual $a_i$. Alternatively, the set
${\cal A}$ may remain (possibly partially) unknown, such that author
disambiguation boils down to finding a partition ${\cal C}$ where
subsets $c_i$  each corresponds to the set of all signatures from a same
individual. Finally, in the case of partially annotated databases as studied in
this work, the setting extends with the partial knowledge of pairs $(a, s) \in
{\cal A} \times {\cal S}$, each indicating that signature $s$ is known to
belong to individual $a$.

>>>>

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem, digital
libraries are increasingly offering tools for authors to manually curate their
publications and claim which ones are theirs. Indirectly, these tools allow for
the inexpensive collection of large annotated training data, which can be
further leveraged to build an automated disambiguation system capable of
inferring patterns for identifying publications written by a same
person.  In this context, we propose a semi-supervised machine learning
disambiguation solution in the case large and partially annotated databases of
scientific publications. Using a real-world dataset of 1 million manual
annotations, we demonstrate how we exploited this data (i) to learn an accurate
classifier for identifying corefering authors, (ii) to guide the clustering of
scientific publications by distinct authors in a semi-supervised way and
finally (iii) to match the resulting clusters with existing profiles.  Our
analysis on this large training set reveal that...
\glnote{To be completed with actual results.}

@@

In academic digital libraries, author name disambiguation is the problem of
grouping together publications written by a same person.  Author name
disambiguation is often a difficult problem because an author may use different
spellings, variants or synonyms across her career and/or distinct authors may
share the same name. Most notably, author disambiguation is often even more
troublesome in the case of researchers from non-Western cultures, where
personal names may be traditionally less diverse (leading to homonym issues) or
for which transliteration to latin characters may not be unique (leading to
synonym issues). With the fast growth of the scientific literature, author
disambiguation has become a pressing issue because the accuracy of information
managed at the level of individuals directly affects the relevance search
results (e.g., when querying for all publications written by a given author),
the reliability of bibliometrics (e.g., citation counts or other impact
metrics) or the relevance of scientific network analysis.

@@

Formally, let us assume a set of publications ${\cal P} = \{ p_0, ...,
p_{N-1}\}$ along with the set of unique individuals ${\cal A} = \{ a_0, ...,
a_{M-1}\}$ having together authored all publications in ${\cal P}$.  Let us
define a signature $s \in p$ from a publication as a unique piece of
information identifying one of the authors of $p$ (e.g., the author name, his
affiliation, along with any other metadata that can be derived from $p$). Let us
denote by ${\cal S} = \{ s | s \in p, p \in {\cal P} \}$ the set of all
signatures that can be extracted from all publications in ${\cal P}$. In this
framework, author disambiguation can be stated as the problem of
finding a partition ${\cal C} = \{ c_0, ..., c_{M-1} \}$ of ${\cal S}$ such
that ${\cal S} = \cup_{i=0}^{M-1} c_i$, $c_i \cap c_j = \phi$ for all $i \neq
j$, and where subsets $c_i$, or clusters, each corresponds to the set of all
signatures belonging to the same individual $a_i$. Alternatively, the set
${\cal A}$ may remain (possibly partially) unknown, such that author
disambiguation boils down to finding a partition ${\cal C}$ where
subsets $c_i$  each corresponds to the set of all signatures from a same
individual. Finally, in the case of partially annotated databases as studied in
this work, the setting extends with the partial knowledge of pairs $(a, s) \in
{\cal A} \times {\cal S}$, each indicating that signature $s$ is known to
belong to individual $a$.
